### 2018-06-15 paper/5-implementations.tex
Work on S5

In reality \Excel works slightly differently (as far as we are able to ascertain) --
on a change it propagates the dirty bits forward using the \hs{dynamicDependencies}, then only checks if the current \hs{key} is dirty. While both methods are equivalent, merely changing the interleaving, our approach allows us to model more of the behaviour of \Excel.

@@

The \Shake approach for dependency tracking involves recording traces and verifying them, for which we use the \hs{Trace} type defined in \ref{sec-verifying-traces}. Complete code is given in \S\ref{fig-shake-implementation}, split into three functions:

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
bazel :: (Eq k, Hashable v) => Build Applicative (Traces k v) k v
bazel = topological $ \key ds act -> do
    s <- get
    let Traces traces contents = getInfo s
    poss <- traceMatch (\k v -> return $ getHash k s == v) key traces
    if null poss then do
        v <- act
        modify $ \s ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
                ts = Traces (t : traces) (Map.insert (hash v) v contents)
            in putInfo ts (putValue key v s)
    else do
        when (getHash key s `notElem` poss) $
            modify $ putValue key (contents Map.! head poss)
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Bazel using our framework; \hs{topological} is
defined in Fig.~\ref{fig-make-implementation}.}\label{fig-bazel-implementation}
\vspace{-2mm}
\end{figure}

Now we have seen all three dependency schemes, we can directly reuse \hs{topological} to define \Bazel. Furthermore, as \Bazel is a tracing build system, we can reuse \hs{Trace} and \hs{traceMatch}, along with the \hs{Traces} type from \S\ref{sec-constructive-traces}. With these pieces in place, the implementation of \Bazel is given in Fig.~\ref{fig-bazel-implementation}. We first figure out the possible results given the current state. If there are no recorded traces for the current dependencies we run it and record the results, otherwise grab a suitable result from the \hs{contents} cache.

@@

Using the abstractions and approaches built thus far, we have shown how to combine dependency scheme and change approach to reproduce existing build systems. In the attached materials we have implemented 9~build systems corresponding to all three dependency schemes, matched with all three change approaches. To us, the most interesting build system as yet unavailable would matching recursive ordering with constructive traces~--~providing a cloud-capable build system with minimality, cutoff and monadic dependencies. Using our framework it is possible to define and test such a system as per Fig.~\ref{fig-cloudshake-implementation}.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
cloudShake :: (Eq k, Hashable v) => Build Monad (Traces k v) k v
cloudShake = recursive $ \key fetch act -> do
    s <- gets fst
    let Traces traces contents = getInfo s
    poss <- traceMatch (\k v -> (==) v . hash <$> fetch k) key traces
    if null poss then do
        (v, ds) <- act
        modify $ \(s,done) ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
                ts = Traces (t : traces) (Map.insert (hash v) v contents)
            in (putInfo ts (putValue key v s), done)
    else do
        s <- gets fst
        when (getHash key s `notElem` poss) $
            modify $ \(s, done) -> (putValue key (contents Map.! head poss) s, done)
\end{minted}
\vspace{-2mm}
\caption{An implementation of Cloud \Shake using our framework.}\label{fig-cloudshake-implementation}
\vspace{-4mm}
\end{figure}

@@

>>>>

In reality \Excel works slightly differently (as far as we are able to
ascertain)~--~on a change it propagates the dirty bits forward using the
\hs{dynamicDependencies}, then only checks if the current \hs{key} is dirty.
While both methods are equivalent, merely changing the interleaving, our
approach allows us to model more of the behaviour of \Excel.

@@

The \Shake approach for dependency tracking involves recording traces and
verifying them, for which we use the \hs{Trace} type defined
in~\ref{sec-verifying-traces}. Complete code is given
in~\S\ref{fig-shake-implementation}, split into three functions:

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Bazel build system, implementation of 'restarting2' is omitted (22 lines)
bazel :: (Ord k, Hashable v) => Build Monad (CT k v) k v
bazel = restarting2 ctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Cloud Shake build system, implementation of 'recursive' is given in Fig. 8
cloudShake :: (Ord k, Hashable v) => Build Monad (CT k v) k v
cloudShake = recursive ctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on constructive traces
ctRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (CT k v) k v
ctRebuilder key value task = Task $ \fetch -> do
    ct <- get
    maybeCachedValue <- constructCT key value (fmap hash . fetch) ct
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- track task fetch
            put =<< recordCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Bazel and Cloud \Shake using our framework;
\hs{restarting2} differs from the restarting scheduler used by Excel
(Fig.~\ref{fig-excel-implementation}) in that it does not rely on the
calculation chain.}
% \hs{topological} is defined in Fig.~\ref{fig-make-implementation}.
\label{fig-bazel-cloud-shake-implementation}
\vspace{-2mm}
\end{figure}

Now we have seen all three dependency schemes, we can directly reuse
\hs{topological} to define \Bazel. Furthermore, as \Bazel is a tracing build
system, we can reuse \hs{Trace} and \hs{traceMatch}, along with the \hs{Traces}
type from \S\ref{sec-constructive-traces}. With these pieces in place, the
implementation of \Bazel is given in
Fig.~\ref{fig-bazel-cloud-shake-implementation}. We first figure out the
possible results given the current state. If there are no recorded traces for
the current dependencies we run it and record the results, otherwise grab a
suitable result from the \hs{contents} cache.

@@

Using the abstractions and approaches built thus far, we have shown how to
combine dependency scheme and change approach to reproduce existing build
systems. In the attached materials we have implemented 9~build systems
corresponding to all three dependency schemes, matched with all three change
approaches. To us, the most interesting build system as yet unavailable would
matching recursive ordering with constructive traces~--~providing a cloud-capable
build system with minimality, cutoff and monadic dependencies. Using our
framework it is possible to define and test such a system as per
Fig.~\ref{fig-bazel-cloud-shake-implementation}.

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Buck build system, implementation of 'topological' is given in Fig. 6
buck :: (Hashable k, Hashable v) => Build Applicative (DCT k v) k v
buck = topological (unliftRebuilder dctRebuilder)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a monadic rebuilder to the corresponding applicative one
unliftRebuilder :: Rebuilder Monad i k v -> Rebuilder Applicative i k v
unliftRebuilder rebuilder key value task = rebuilder key value $ Task $ run task
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Nix build system, implementation of 'recursive' is given in Fig. 8
nix :: (Hashable k, Hashable v) => Build Monad (DCT k v) k v
nix = recursive dctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on deterministic constructive traces
dctRebuilder :: (Hashable k, Hashable v) => Rebuilder Monad (DCT k v) k v
dctRebuilder key _value task = Task $ \fetch -> do
    dct <- get
    maybeCachedValue <- constructDCT key (fmap hash . fetch) dct
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- trackM task fetch
            put =<< recordDCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Buck and \Nix using our framework.}
\label{fig-buck-nix-implementation}
\vspace{-2mm}
\end{figure}
